A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution

### 2019-07-26 paper/journal.tex
Most of the comments have been addressed.

In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing. How would one increase the crossing resolution of a given drawing? First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase. Of course, we cannot consider all possible positions for the vertex to be moved. Instead, we consider a small set of randomly generated ones. If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position.

Note that, in general, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs of general graphs. Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we followed a more practical approach. We implemented our algorithm and the force-directed ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs that are widely used in Graph Drawing for comparing different drawing algorithms. Our evaluation indicates that our method significantly outperforms the aforementioned force-directed algorithms in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs. Analogous results are obtained, when our algorithm and the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize the \emph{angular resolution} (that is, the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (that is, the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).

@@

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2017; the last two columns correspond to results presented in this paper (with and without time limitations, respectively).}
\centering
\begin{adjustbox}{width=1\textwidth}
\begin{tabular}{c|c|c|c|c|c|c}
\toprule
\rowcolor{gray!30!}~Graph~ & ~no.~of vertices~ & ~no.~of edges~ & CoffeeVM & T\"ubingenMidnight & Time restricted & No-time restriction \\
\midrule
\rowcolor{gray!30!} 1  &  9  & 18    & 90$^\circ$ & 77$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!00!} 2  & 12  & 36    & 88$^\circ$ & 42$^\circ$ & 88$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 3  & 16  & 42    & 90$^\circ$ & 89$^\circ$ & 88$^\circ$ & 90$^\circ$\\
\rowcolor{gray!00!} 4  & 17  & 54    & 89$^\circ$ & 89$^\circ$ & 77$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 5  & 30  & 95    & 80$^\circ$ & 30$^\circ$ & 79$^\circ$ & 87$^\circ$\\
\rowcolor{gray!00!} 6  & 40  & 75    & 90$^\circ$ & 78$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!30!} 7  & 80  & 254   & 57$^\circ$ & 34$^\circ$ & 56$^\circ$ & 64$^\circ$\\
\rowcolor{gray!00!} 8  & 85  & 136   & 85$^\circ$ & 61$^\circ$ & 81$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 9  & 149 & 258   & 60$^\circ$ & 9$^\circ$  & 55$^\circ$ & 88$^\circ$\\
\rowcolor{gray!00!} 10 & 174 & 344   & 21$^\circ$ & 4$^\circ$  & 24$^\circ$ & 24$^\circ$\\
\rowcolor{gray!30!} 11 & 250 & 400   & 47$^\circ$ & 6$^\circ$  & 57$^\circ$ & 72$^\circ$\\
\rowcolor{gray!00!} 12 & 300 & 987   & 36$^\circ$ & 5$^\circ$  & 26$^\circ$ & 36$^\circ$\\
\rowcolor{gray!30!} 13 & 392 & 562   & 25$^\circ$ & 4$^\circ$  & 22$^\circ$ & 34$^\circ$\\
\rowcolor{gray!00!} 14 & 510 & 820   & 34$^\circ$ & 5$^\circ$  & 30$^\circ$ & 43$^\circ$\\
\rowcolor{gray!30!} 15 & 650 & 2,173 & 21$^\circ$ & 4$^\circ$  & 13$^\circ$ & 29$^\circ$\\
\rowcolor{gray!00!} 16 & 998 & 1,596 & 20$^\circ$ & 3$^\circ$  & 12$^\circ$ & 26$^\circ$\\
\bottomrule
\end{tabular}
\end{adjustbox}
\label{tab:gdContest2017}
\end{table*}

@@

%\myparagraph{Graph Drawing Contest 2017} 
In 2017, we participated to the contest (by implementing a mixture of the two force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}). 
%
Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.

It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''; details for their approach can be found in~\cite{DBLP:journals/corr/abs-1807-09483}. Our team ``\texttt{TuebingenMidnight}''  performed miserably; we actually took the third position out of a total of four teams that participated. As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$. In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as s results, which were achieved without any time limitations (see column ``No-time restriction'').

@@

%\myparagraph{Graph Drawing Contest 2018} 
In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}'' (using an implementation of the algorithm presented in this paper). The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''. Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.

@@

Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of them to maximize exclusively the corresponding measures. In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we discussed in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution). The reported results, illustrated in Figures~\ref{fig:cr-res}--\ref{fig:northAngular}, are on average across different drawings with same number of vertices.

@@

\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}. Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution). It is immediate to see that our algorithm outperforms all other ones in terms of the crossing resolution of the produced drawings, when we do not impose any restriction on the edge-length ratio of the computed drawings; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}. The variant of our algorithm, which does not violate the edge-length ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ELR-restricted}, in Fig.~\ref{fig:cr-res-1}. Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).

@@

While our unrestricted algorithm produces drawings with better crossing resolution, this comes at a cost of drastically increased edge-length ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding edge-length ratio of the drawings produced by the algorithm of Argyriou et al. For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short. Going back to our unrestricted algorithm, its behavior is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose sping forces inpose stong restrictions on the edge lengths). On the other hand, the restricted variant of our algorithm, which does not allow the edge-length ratio to increase, has more or less comparable performance (in terms of edge-length ratio) as the one of Huang et al.

Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings, which at the same time is significantly smaller than the  number of crossings produced by the two other algorithms; see Fig.~\ref{fig:cr-res-3}.

A different behavior can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}. We note here that we used different criteria to determine whether the algorithms of our experiment had converged. For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees. For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations. We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment. In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees. Observe that even under this different convergence criterion, the algorithm needs significantly more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.

@@

\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}. Here, each algorithm was adjusted to maximize both the crossing and the angular resolution, simultaneously. For the vast majority of the graphs in the experiment, both our unrestricted algorithm and its restricted variant yield drawings with better total resolution than the corresponding ones by Argyriou et al. The drawings produced by the algorithm by Huang et al.\ seems to have worse total resolution; see Fig.~\ref{fig:to-res-1}. Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.

@@

\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the results for the angular resolution; see Fig.~\ref{fig:an-res}. Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution). A notable observation is that, for small graphs the best results are achieved by the algorithm by Argyriou et al., while for medium-size graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}. For large graphs, the two algorithms tend to have the same performance. The restricted variant of our algorithm yields drawings with slightly worse angular resolution. The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.

The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}. This observation suggests that, for most of the graphs of our experiment, the angular resolution dominates the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results. The small differences result from the fact that the crossing resolution cannot be entirely neglected.

>>>>

In this work, we approach the crossing resolution maximization problem from a different perspective. We suggest a simple and intuitive randomization method for computing drawings with high crossing resolution, which, in a sense, mimics the way a human would try to increase the crossing resolution of a drawing. How would one increase the crossing resolution of a given drawing? First, she would try to identify the pair of edges that define the crossing resolution of the drawing (we call them \emph{critical} edges); then, she would try to move an endvertex of this pair (which we choose at random), hoping that by this move the crossing resolution will increase. Of course, we cannot consider all possible positions for the vertex to be moved. Instead, we consider a small set of randomly generated ones. If there exists a position among them, that does not lead to a reduction of the crossing resolution, we move the vertex to this position. Note at that in general the vertex movement paradigm is an established technique and has successfully applied to graph theoretic and geometric problems; see, e.g.,~\cite{DBLP:conf/sofsem/BlasiusRR17,DBLP:conf/dac/FiducciaM82,6771089,DBLP:conf/alenex/RadermacherRRW18,DBLP:conf/wsdm/UganderB13}.

On the other hand, randomization is a technique that has not been deeply examined in Graph Drawing, as it seems difficult to even speculate about the expected quality of the produced drawings; a notable exception is the randomized approach by Goldschmidt and Takvorian~\cite{DBLP:journals/networks/GoldschmidtT94} for computing large planar subgraphs of general graphs. Since we also could not provide any theoretical guarantee on the expected quality of the produced drawings (mainly due to the nature of the problem itself), we followed a more practical approach. We implemented our algorithm and the force-directed ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, and we experimentally compared them on standard benchmark graphs that are widely used in Graph Drawing for comparing different drawing algorithms. Our evaluation indicates that our method significantly outperforms the aforementioned force-directed algorithms in terms of crossing resolution, but this comes at the cost of slightly worse running time, especially for large and dense graphs. Analogous results are obtained, when our algorithm and the ones by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} are adjusted to maximize the \emph{angular resolution} (that is, the minimum value of the angle between any two adjacent edges~\cite{DBLP:journals/siamcomp/FormannHHKLSWW93}) or the \emph{total resolution} (that is, the minimum of the angular and the crossing resolution~\cite{DBLP:journals/cj/ArgyriouBS13}).

@@

\begin{table*}
\caption{Summary of the results for the Graph Drawing Contest 2017; the last two columns correspond to results obtained by the implementation of the algorithm presented in this paper (with and without time limitations, respectively).}
\centering
\begin{adjustbox}{width=1\textwidth}
\begin{tabular}{c|c|c|c|c|c|c}
\toprule
\rowcolor{gray!30!}~Graph~ & ~no.~of vertices~ & ~no.~of edges~ & CoffeeVM & T\"ubingenMidnight & Time restricted & No-time restriction \\
\midrule
\rowcolor{gray!30!} 1  &  9  & 18    & 90$^\circ$ & 77$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!00!} 2  & 12  & 36    & 88$^\circ$ & 42$^\circ$ & 88$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 3  & 16  & 42    & 90$^\circ$ & 89$^\circ$ & 88$^\circ$ & 90$^\circ$\\
\rowcolor{gray!00!} 4  & 17  & 54    & 89$^\circ$ & 89$^\circ$ & 77$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 5  & 30  & 95    & 80$^\circ$ & 30$^\circ$ & 79$^\circ$ & 87$^\circ$\\
\rowcolor{gray!00!} 6  & 40  & 75    & 90$^\circ$ & 78$^\circ$ & 90$^\circ$ & 90$^\circ$\\
\rowcolor{gray!30!} 7  & 80  & 254   & 57$^\circ$ & 34$^\circ$ & 56$^\circ$ & 64$^\circ$\\
\rowcolor{gray!00!} 8  & 85  & 136   & 85$^\circ$ & 61$^\circ$ & 81$^\circ$ & 89$^\circ$\\
\rowcolor{gray!30!} 9  & 149 & 258   & 60$^\circ$ & 9$^\circ$  & 55$^\circ$ & 88$^\circ$\\
\rowcolor{gray!00!} 10 & 174 & 344   & 21$^\circ$ & 4$^\circ$  & 24$^\circ$ & 24$^\circ$\\
\rowcolor{gray!30!} 11 & 250 & 400   & 47$^\circ$ & 6$^\circ$  & 57$^\circ$ & 72$^\circ$\\
\rowcolor{gray!00!} 12 & 300 & 987   & 36$^\circ$ & 5$^\circ$  & 26$^\circ$ & 36$^\circ$\\
\rowcolor{gray!30!} 13 & 392 & 562   & 25$^\circ$ & 4$^\circ$  & 22$^\circ$ & 34$^\circ$\\
\rowcolor{gray!00!} 14 & 510 & 820   & 34$^\circ$ & 5$^\circ$  & 30$^\circ$ & 43$^\circ$\\
\rowcolor{gray!30!} 15 & 650 & 2,173 & 21$^\circ$ & 4$^\circ$  & 13$^\circ$ & 29$^\circ$\\
\rowcolor{gray!00!} 16 & 998 & 1,596 & 20$^\circ$ & 3$^\circ$  & 12$^\circ$ & 26$^\circ$\\
\bottomrule
\end{tabular}
\end{adjustbox}
\label{tab:gdContest2017}
\end{table*}

@@

%\myparagraph{Graph Drawing Contest 2017} 
In 2017, we participated to the contest~\cite{DBLP:conf/gd/DevannyKLR17} (by implementing a mixture of the two force-directed algorithms by Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and by Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}). 
%
Table~\ref{tab:gdContest2017} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2017/results.html}.

It is far from clear that the wining team was the team  ``\texttt{CoffeeVM}''. Notably,  their approach is in principle very similar to the one that we present in this paper. In particular, they also adopt the vertex movement paradigm on a set of a randomly chosen samples. Once they find a sample position, which improves the crossing resolution of the drawing, then they focus on a region close to this position. In this region, they are examining more sample positions, hoping that they will find one that will further improve the crossing resolution of the drawing; for details refer to~\cite{DBLP:conf/gd/DemelDMRW18}.

Our team ``\texttt{TuebingenMidnight}''  performed miserably; we actually took the third position out of a total of four teams that participated. As already mentioned, each team had only one hour to compute a layout for each of the challenge graphs, which in year 2017 were in total $16$. In Table~\ref{tab:gdContest2017}, we additionally provide results that were achieved by the algorithm presented in this paper in two settings; within the time limit of the contest (see column ``Time restricted''), as well as  results, which were achieved without any time limitations (see column ``No-time restriction'').

@@

%\myparagraph{Graph Drawing Contest 2018} 
In 2018, we participated again to the Graph Drawing contest under the name ``\texttt{T\"ubingenColdShower}''~\cite{DBLP:conf/gd/DevannyKLR18} (using an implementation of the algorithm presented in this paper). The winning team of the Graph Drawing contest 2017 also participated under the name ``\texttt{CoffeeVM+}''. Table~\ref{tab:gdContest2018} summarizes the results, which are also available at \texttt{www.graphdrawing.de/contest2018/results.html}.

@@

Since all algorithms of the experiment can easily be adjusted to maximize~only the crossing resolution, or only the angular resolution or both (by maximizing~the total resolution), we adjusted each of them to maximize exclusively the corresponding measures. In our algorithm, this can be achieved by modifying appropriately the content of the vertex-pool (as we discussed in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.\ and of Huang et al.\ by switching on only the forces that maximize the corresponding properties under measure (note that, each of these two algorithms has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution). The reported results, illustrated in Figures~\ref{fig:cr-res}--\ref{fig:northAngular}, are on average across different drawings with same number of vertices on the following set of parameters:
%
\begin{enumerate}[P.1.]
\item \label{p:cr-res} crossing resolution
\item \label{p:to-res} total resolution
\item \label{p:an-res} angular resolution
\item \label{p:as-rat} edge-length ratio
\item \label{p:no-xing} number of crossings
\end{enumerate}

@@

\myparagraph{Crossing resolution} Our results for the crossing resolution are summarized in Fig.~\ref{fig:cr-res}. Here, each algorithm was adjusted to maximize exclusively the crossing resolution (i.e., by ignoring the drawing's angular resolution). It is immediate to see that on average the crossing resolution of the drawings obtained by your algorithm, which does not impose any restriction on the edge-length ratio of the computed drawings, is considerably larger than the crossing resolutions of the remaining algorithms; refer to the solid-black curve, denoted as \emph{Unrestricted}, in Fig.~\ref{fig:cr-res-1}. The variant of our algorithm, which does not violate the edge-length ratio of the initial layout, leads to drawings with slightly smaller crossing resolution; refer to the solid-gray curve, denoted as \emph{ELR-restricted}, in Fig.~\ref{fig:cr-res-1}. Finally, the two force-directed algorithms seem to produce drawings with worse crossing resolution; refer to the dotted-gray and dotted-black curves of Fig.~\ref{fig:cr-res-1} (by Argyriou et al.\ and by Huang et al., respectively).

@@

While our unrestricted algorithm produces drawings with larger crossing resolution on average, this comes at a cost of drastically increased edge-length ratio (see Fig.~\ref{fig:cr-res-2}), which, however, is still better that the corresponding edge-length ratio of the drawings produced by the algorithm of Argyriou et al. For the latter algorithm, it seems that the forces due to the angles formed at the crossings outperform the corresponding spring forces, which try to keep the lengths of the edges short. Going back to our unrestricted algorithm, its behavior is up to a certain degree expected, mainly due to the fact that there is no control on the lengths of the edges (as it is, e.g., in the case of the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}, whose spring forces impose strong restrictions on the edge lengths). On the other hand, the edge-length ratio of the drawings obtained by the restricted variant of our algorithm, which does not allow the edge-length ratio to increase, is very close to the edge-length ratio of the drawings obtained by the algorithm of Huang et al. Note that the latter algorithm is the one yielding drawings that on average have the lowest edge-length ratio. Regarding the number of crossings, the restricted variant of our algorithm and the force-directed algorithm of Huang et al.\ yield drawings with comparable number of crossings. On the other hand, the two other algorithms of our experiment produce drawings that on average have more crossings; see Fig.~\ref{fig:cr-res-3}.

A different behavior can be observed in the number of iterations, which are required by the algorithms to converge; refer to Fig.~\ref{fig:cr-res-4}. We note here that we used different criteria to determine whether the algorithms of our experiment had converged. For our algorithms and for the force-directed algorithm by Huang et al., we assumed that the algorithm had converged, if the crossing resolution between 500 consecutive iterations was not improved by more than 0.001 degrees. For the algorithm by Argyriou et al., we decided to use a much more restricted convergence criterion, because the produced layouts can change vastly between  consecutive iterations. We made this choice mainly to have ``comparable'' number of iterations among the algorithms of the experiment. In this direction, we adopted the convergence criterion that the authors used in their previous experimental analysis, that is, we assumed that the algorithm had converged, if the crossing resolution between two consecutive iterations was not improved by more than 0.001 degrees. Observe that even under this different convergence criterion, the algorithm needs more iterations to converge than the remaining three algorithms of the experiment; see Fig.~\ref{fig:cr-res-4}.

@@

\myparagraph{Total resolution} Our results for the total resolution are summarized in Fig.~\ref{fig:to-res}. Here, each algorithm was adjusted to maximize both the crossing and the angular resolution, simultaneously. For graphs with more 20 vertices, both our unrestricted algorithm and its restricted variant yield drawings with better total resolution than the corresponding ones by Argyriou et al. The drawings produced by the algorithm by Huang et al.\ seem to have worse total resolution; see Fig.~\ref{fig:to-res-1}.%Note, however, that both variants of our algorithm as well as the force-directed algorithm by Argyriou et al.\ tend to produce drawings of the same total resolution for larger graphs with a small difference in our favor.

@@

\myparagraph{Angular resolution} We conclude the analysis of our experimental evaluation with the results for the angular resolution; see Fig.~\ref{fig:an-res}. Here, each algorithm was adjusted  to maximize only the angular resolution (i.e., by ignoring the drawing's crossing resolution). A notable observation is that, for graphs with less than $30$ vertices the best results are achieved by the algorithm by Argyriou et al., while for the remaining graphs by our unrestricted algorithm; see Fig.~\ref{fig:an-res-1}. For large graphs, the two algorithms tend to have the same performance. The restricted variant of our algorithm yields drawings with slightly worse angular resolution. The algorithm by Huang et al.\ is outperformed by all algorithms of the experiment.
%
The results for the edge-length ratio, the number of crossings and the required number of iterations are very similar with corresponding ones for the total resolution; see Figs.~\ref{fig:an-res-2}--\ref{fig:an-res-4}.% This observation suggests that, for most of the graphs of our experiment, the angular resolution is greater than the crossing resolution (and thus is the one defining the total resolution) in the constructed drawings, which explains the similarity in the reported results. The small differences result from the fact that the crossing resolution cannot be entirely neglected.
